-------------------
title: "DATS-6101"
author: "GROUP 7"
date: "2023-10-12"
output: 
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
--------------------
Our Project uses a Apple Appstore data which has 21 columns. The variables are:
App_Id
App_Name
AppStore_Url
Primary_Genre
Content_Rating
Size_Bytes
Required_IOS_Version
Released
Updated
Version
Price
Currency
Free
DeveloperId
Developer
Developer_Url
Developer_Website
Average_User_Rating
Reviews
Current_Version_Score
Current_Version_Reviews

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ezids)
```
#Reading the Apple App Data CSV into the data frame.
```{r}
appData = read.csv('appleAppData.csv')
```
#Dimenstions of the data frame.
```{r}
dim(appData)
```
#Summary of the dataset
```{r}
summary(appData)
```
#First five and Last Five observations of the data frame
```{r}
head(appData,5)
tail(appData,5)
```
#Listing the numerical variables(columns)
```{r}
numerical_cols <- sapply(appData, is.numeric)
numerical_data <- appData[, numerical_cols]
numerical_data
```
#Check for missing values in the datset
```{r}

```

#A barplot showing the number of missing values
#Removing the attribute having the highest missing values.


#Removing the NA value in the column AppNAme
```{r}
appData <- subset(appData, !is.na(App_Name))
```

