---
title: "DATS-6101"
author: "GROUP 7"
date: "2023-10-12"
output: 
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---
Our Project uses a Apple Appstore data which has 21 columns.

The variables are:
App_Id
App_Name
AppStore_Url
Primary_Genre
Content_Rating
Size_Bytes
Required_IOS_Version
Released
Updated
Version
Price
Currency
Free
DeveloperId
Developer
Developer_Url
Developer_Website
Average_User_Rating
Reviews
Current_Version_Score
Current_Version_Reviews

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ezids)
library(tidyr)
```

### Reading the Apple App Data CSV into the data frame

```{r}
appData = read.csv('appleAppData.csv')
```

### Dimenstions of the data frame

```{r}
dim(appData)
```
### Summary of the dataset

```{r}
summary(appData)
```
### First five and Last five observations of the data frame

```{r}
head(appData,5)
tail(appData,5)
```
### Listing the numerical variables(columns)

```{r}
numerical_cols <- sapply(appData, is.numeric)
numerical_data <- appData[, numerical_cols]
# numerical_data
```

### Removing the NA value in the column AppNAme

```{r}
appData <- subset(appData, !is.na(App_Name))
```

### Check for missing values in the datset

```{r}
missing_values <- sum(is.na(appData))
cat("Total missing values in the dataset:", missing_values, "\n")
missing_values_per_column <- colSums(is.na(appData))
cat("Missing values per column:\n")
print(missing_values_per_column)
```

```{r}
# Count the number of blank values in Developer_Url
developer_url_column <- appData$Developer_Url

blank_columns <- sum(is.na(developer_url_column) | developer_url_column == "")
cat("Number of blank columns in 'Developer_Url':", blank_columns, "\n")
```


```{r}
# Count the number of blank values in Developer_Website
developer_website_column <- appData$Developer_Website

blank_values <- sum(is.na(developer_website_column) | developer_website_column == "")
cat("Number of blank values in 'Developer_Website':", blank_values, "\n")
```

### Checking all missing values in the dataset
```{r}
# Assuming 'appData' is your dataset

# Total missing values in the dataset
missing_values <- sum(is.na(appData))
cat("Total missing values in the dataset:", missing_values, "\n")

# Missing values per column
missing_values_per_column <- colSums(is.na(appData))
cat("Missing values per column:\n")
print(missing_values_per_column)

# Number of missing values in 'Developer_Url' column
blank_developer_url <- sum(is.na(appData$Developer_Url) | appData$Developer_Url == "")
cat("Number of missing values in 'Developer_Url' column:", blank_developer_url, "\n")

# Number of missing values in 'Developer_Website' column
blank_developer_website <- sum(is.na(appData$Developer_Website) | appData$Developer_Website == "")
cat("Number of missing values in 'Developer_Website' column:", blank_developer_website, "\n")

```

### Barplot showing the number of missing values
```{r}
# Load required library
library(ggplot2)

# Calculate missing values per column
missing_values_per_column <- colSums(is.na(appData))

# Create a data frame with column names and missing values
missing_data <- data.frame(Column = names(missing_values_per_column),
                           MissingValues = missing_values_per_column)

ggplot(missing_data, aes(x = Column, y = MissingValues)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Column Names", y = "Number of Missing Values", 
       title = "Number of Missing Values in the dataset")

```


```{r}
# Calculate the number of missing values in 'Developer_Url' and 'Developer_Website' columns
blank_developer_url <- sum(is.na(appData$Developer_Url) | appData$Developer_Url == "")
blank_developer_website <- sum(is.na(appData$Developer_Website) | appData$Developer_Website == "")

# Create a data frame for plotting
missing_data <- data.frame(
  Column = c('Developer_Url', 'Developer_Website'),
  Missing_Values = c(blank_developer_url, blank_developer_website)
)


# Create a bar plot
ggplot(missing_data, aes(x = Column, y = Missing_Values, fill = Column)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of Missing Values in Columns",
       x = "Column Names",
       y = "Number of Missing Values") +
  theme_minimal()

```
### Dropping Developer_Website column

```{r}
# Dropping Developer_Website column
appData = subset(appData, select = -c(Developer_Website))
head(appData, 1)
```

### Updating Price and Size_Bytes columns' NA values with mean

```{r}
df_p_sb <- subset(appData, select=c("Price", "Size_Bytes"))
head(df_p_sb, 1)
meanvals <- as.list(colMeans(df_p_sb, na.rm = TRUE))
print(meanvals)
df_p_sb <- replace_na(df_p_sb, meanvals)
head(df_p_sb, 1)
appData = subset(appData, select = -c(Price))
appData = subset(appData, select = -c(Size_Bytes))
appData <- cbind(appData, Price = df_p_sb$Price)
appData <- cbind(appData, Size_Bytes = df_p_sb$Size_Bytes)
head(appData, 1)
```
```{r}
# write.csv(appData, "C:\\Users\\Krishnasurya\\Documents\\exported\\appData.csv", row.names=TRUE)
# Calculate missing values per column
missing_values_per_column_2 <- colSums(is.na(appData))

# Create a data frame with column names and missing values
missing_data_2 <- data.frame(Column = names(missing_values_per_column_2),
                           MissingValues = missing_values_per_column_2)

ggplot(missing_data_2, aes(x = Column, y = MissingValues)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Column Names", y = "Number of Missing Values 2", 
       title = "Number of Missing Values in the dataset")

missing_values_per_column <- colSums(is.na(appData))
print(missing_values_per_column <- colSums(is.na(appData)))

# appData = appData[appData$App_Name != 'nan', ]

# # Calculate missing values per column
# missing_values_per_column_3 <- colSums(is.na(appData))
# 
# # Create a data frame with column names and missing values
# missing_data_3 <- data.frame(Column = names(missing_values_per_column_3),
#                            MissingValues = missing_values_per_column_3)
# 
# ggplot(missing_data_3, aes(x = Column, y = MissingValues)) +
#   geom_bar(stat = "identity", fill = "skyblue") +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
#   labs(x = "Column Names", y = "Number of Missing Values 3",
#        title = "Number of Missing Values in the dataset")
```
### Number of duplicates in the dataset

```{r}
print(sum(duplicated(appData)))
```