-------------------
title: "DATS-6101"
author: "GROUP 7"
date: "2023-10-12"
output: 
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
--------------------
Our Project uses a Apple Appstore data which has 21 columns. The variables are:
App_Id
App_Name
AppStore_Url
Primary_Genre
Content_Rating
Size_Bytes
Required_IOS_Version
Released
Updated
Version
Price
Currency
Free
DeveloperId
Developer
Developer_Url
Developer_Website
Average_User_Rating
Reviews
Current_Version_Score
Current_Version_Reviews

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ezids)
```
#Reading the Apple App Data CSV into the data frame.
```{r}
appData = read.csv('appleAppData.csv')
```
#Dimenstions of the data frame.
```{r}
dim(appData)
```
#Summary of the dataset
```{r}
summary(appData)
```
#First five and Last five observations of the data frame
```{r}
head(appData,5)
tail(appData,5)
```
#Listing the numerical variables(columns)
```{r}
numerical_cols <- sapply(appData, is.numeric)
numerical_data <- appData[, numerical_cols]
numerical_data
```
#Check for missing values in the datset
```{r}

```

#Removing the NA value in the column AppNAme
```{r}
#appData <- subset(appData, !is.na(App_Name))
```

```{r}
missing_values <- sum(is.na(appData))
cat("Total missing values in the dataset:", missing_values, "\n")
missing_values_per_column <- colSums(is.na(appData))
cat("Missing values per column:\n")
print(missing_values_per_column)
```

```{r}
# Count the number of blank values in Developer_Url
developer_url_column <- appData$Developer_Url

blank_columns <- sum(is.na(developer_url_column) | developer_url_column == "")
cat("Number of blank columns in 'Developer_Url':", blank_columns, "\n")
```


```{r}
# Count the number of blank values in Developer_Website
developer_website_column <- appData$Developer_Website

blank_values <- sum(is.na(developer_website_column) | developer_website_column == "")
cat("Number of blank values in 'Developer_Website':", blank_values, "\n")
```

#A single code to check all missing values in our dataset:
```{r}
# Assuming 'appData' is your dataset

# Total missing values in the dataset
missing_values <- sum(is.na(appData))
cat("Total missing values in the dataset:", missing_values, "\n")

# Missing values per column
missing_values_per_column <- colSums(is.na(appData))
cat("Missing values per column:\n")
print(missing_values_per_column)

# Number of missing values in 'Developer_Url' column
blank_developer_url <- sum(is.na(appData$Developer_Url) | appData$Developer_Url == "")
cat("Number of missing values in 'Developer_Url' column:", blank_developer_url, "\n")

# Number of missing values in 'Developer_Website' column
blank_developer_website <- sum(is.na(appData$Developer_Website) | appData$Developer_Website == "")
cat("Number of missing values in 'Developer_Website' column:", blank_developer_website, "\n")

```

#A barplot showing the number of missing values
```{r}
# Load required library
library(ggplot2)

# Calculate missing values per column
missing_values_per_column <- colSums(is.na(appData))

# Create a data frame with column names and missing values
missing_data <- data.frame(Column = names(missing_values_per_column),
                           MissingValues = missing_values_per_column)

ggplot(missing_data, aes(x = Column, y = MissingValues)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Column Names", y = "Number of Missing Values", 
       title = "Number of Missing Values in the dataset")

```


```{r}
# Calculate the number of missing values in 'Developer_Url' and 'Developer_Website' columns
blank_developer_url <- sum(is.na(appData$Developer_Url) | appData$Developer_Url == "")
blank_developer_website <- sum(is.na(appData$Developer_Website) | appData$Developer_Website == "")

# Create a data frame for plotting
missing_data <- data.frame(
  Column = c('Developer_Url', 'Developer_Website'),
  Missing_Values = c(blank_developer_url, blank_developer_website)
)


# Create a bar plot
ggplot(missing_data, aes(x = Column, y = Missing_Values, fill = Column)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of Missing Values in Columns",
       x = "Column Names",
       y = "Number of Missing Values") +
  theme_minimal()

```